
% \documentclass[conference]{IEEEtran}
% \documentclass{sig-alternate} % pre 2017
% \documentclass[sigconf]{acmart}  % starting in 2017
% \documentclass{vldb}
\documentclass[acmlarge]{acmart}
\usepackage{balance}  % vldb requirement; for  \balance command on last page
\input{setup.tex}

\usepackage[all]{nowidow}

% Metadata Information
% \acmJournal{IMWUT}
% \acmVolume{2}
% \acmNumber{3}
% \acmArticle{0}
% \acmYear{2018}
% \acmMonth{9}
% \acmArticleSeq{15}
\setcopyright{acmlicensed}
\acmJournal{IMWUT}
\acmYear{2018} \acmVolume{2} \acmNumber{3} \acmArticle{93} \acmMonth{9} \acmPrice{15.00}\acmDOI{10.1145/3264903}

\received{May 2018}
\received{July 2018}
\received[accepted]{September 2018}

\newcommand{\mine}{\textsc{Sprintz}}
\newcommand{\minesp}{\textsc{Sprintz}\text{ }}
\newcommand{\fire}{\textsc{Fire}\text{ }}
\newcommand{\justfire}{\textsc{Fire}}

% \setcopyright{rightsretained}

\begin{document}

% ================================================================
\title{Sprintz: Time Series Compression for the Internet of Things}
% ================================================================

% \author{Anonymous Authors}
% \affiliation{ \institution{Anonymous Affiliation}}

\author{Davis Blalock}
\affiliation{
  % \institution{Computer Science and Artificial \\ Intelligence Laboratory}
  \institution{Massachusetts Institute of Technology}
}
\email{dblalock@mit.edu}

\author{Samuel Madden}
\affiliation{
  % \institution{Computer Science and Artificial \\ Intelligence Laboratory}
  \institution{Massachusetts Institute of Technology}
}
\email{madden@csail.mit.edu}

\author{John Guttag}
\affiliation{
  % \institution{Computer Science and Artificial \\ Intelligence Laboratory}
  \institution{Massachusetts Institute of Technology}
}
\email{guttag@mit.edu}

% ------------------------------------------------
\begin{abstract}
% ------------------------------------------------

Thanks to the rapid proliferation of connected devices, sensor-generated time series constitute a large and growing portion of the world's data. Often, this data is collected from distributed, resource-constrained devices and centralized at one or more servers. A key challenge in this setup is reducing the size of the transmitted data without sacrificing its quality. Lower quality reduces the data's utility, but smaller size enables both reduced network and storage costs at the servers and reduced power consumption in sensing devices. A natural solution is to compress the data at the sensing devices. Unfortunately, existing compression algorithms either violate the memory and latency constraints common for these devices or, as we show experimentally, perform poorly on sensor-generated time series.

We introduce a time series compression algorithm that achieves state-of-the-art compression ratios while requiring less than 1KB of memory and adding virtually no latency. This method is suitable not only for low-power devices collecting data, but also for servers storing and querying data; in the latter context, it can decompress at over 3GB/s in a single thread, even faster than many algorithms with much lower compression ratios. A key component of our method is a high-speed forecasting algorithm that can be trained online and significantly outperforms alternatives such as delta coding.

Extensive experiments on datasets from many domains show that these results hold not only for sensor data but also across a wide array of other time series.

\end{abstract}

% ------------------------------------------------
% CCS taxonomy stuff / keywords
% ------------------------------------------------

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10002951.10002952.10002971.10003451.10002975</concept_id>
  <concept_desc>Information systems~Data compression</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10002950.10003648.10003688.10003693</concept_id>
  <concept_desc>Mathematics of computing~Time series analysis</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Data compression}
\ccsdesc[300]{Computer systems organization~Embedded systems}
\ccsdesc[100]{Mathematics of computing~Time series analysis}

\keywords{Data Compression, Time Series, Embedded Systems}

\maketitle

\renewcommand{\shortauthors}{D. Blalock et al.}

% ================================================================
\section{Introduction} \label{sec:intro}
% ================================================================

\input{intro.tex}

% % ================================================================
% \section{Definitions and Problem} \label{sec:problem}
% % ================================================================

% \input{problem.tex}

% ================================================================
\vspace{3mm}
\section{Related Work} \label{sec:relatedWork}
% ================================================================

\input{related_work.tex}

% ================================================================
\section{Method} \label{sec:method}
% ================================================================

\input{method.tex}

% ================================================================
\section{Experimental Results} \label{sec:results}
% ================================================================

\input{results.tex}

% % ================================================================
% \vspace{3mm}
% \section{Related Work} \label{sec:relatedWork}
% % ================================================================

% \input{related_work.tex}

% % ================================================================
% \section{Future Work} \label{sec:futureWork}
% % ================================================================

% The above experiments demonstrate that \minesp achieves

% ================================================================
\section{Conclusion} \label{sec:conclusion}
% ================================================================

We introduce \mine, a compression algorithm for multivariate integer time series that achieves state-of-the-art compression ratios across a large number of publicly available datasets. It also attains speeds of up to 3GB/s in a single thread and predictable performance as a function of the number of variables being compressed. Moreover, it only needs to buffer eight samples at a time, enabling low latency for continuously arriving data. Finally, \minesp has extremely low memory requirements, making it feasible to run even on resource-constrained devices.

As part of evaluating \mine, we also conducted what is, to the best of our knowledge, the largest empirical investigation of time series compression that has been reported. To both ensure reproducibility of our work and facilitate future research in this area, we make available all of our experiments as a public benchmark.

In future work, we hope to characterize the relationship between compression and power savings, both for \minesp and for other methods. The savings are upper bounded by the compression ratio in the limit of data transmission consuming all power, but real-world systems have various overheads that cause significant deviation from this idealized model.


% ================================================================
\section*{Acknowledgments}
% ================================================================

This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. 1122374. Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors(s) and do not necessarily reflect the views of the National Science Foundation.

% \IEEEtriggeratref{27}	% trigger column break to make cols even
% \bibliographystyle{ACM-Reference-Format}
\bibliographystyle{abbrv}
\bibliography{doc}

\balance

\end{document}
